---
title: "Project 1-Part 1"
author: "Constantine Sahinidis"
date: "2024-10-04"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculating results

**This function was created  in order to re-use code by changing the following parameters:**

- **Number of weekends**: This parameter represents the number of weeks we run our calculations. We used 4 different timeframes (1 week,1 month, 1 quarter and 1 year) to see both short-term, mid-term and long-term results of our inventory strategies.
- **Order when out**: This parameter is used to enable/disable orders whenever our inventory is empty. After 5 days of each order our inventory is adds 1 scooter.
- **Fixed delivery**: This parameter is used to set a fixed delivery schedule. For example if fixed_Delivery=7 every 7th day of the week we will receive a new scooter.(We make an assumption that orders occur 5 days prior)

## Explaining process occured in the code:

**Variable initialization:**
```{r Explaning calculation function 1,echo=TRUE,eval=FALSE}

my_function <- function(number_of_weekends,
                        order_when_out,
                        fixed_delivery) {
  trials <- 10000 #here we set number of trials
  #number_of_weekends <- This is the timeframe we want to check
  
  #order_when_out   : Flag used for orders in case we are out of stock
  #(If value is 1 then whenever our stock is 0 we make an order)
  
  #fixed_delivery: This parameter is used for deliveries that are on a fixed schedule.
  #If it's 0 the parameter is disabled
  
  lambda <- 1 / 7   #Lambda parameter for Poisson
  
  #To get our final result I implemented a for loop of 10.000 trials and gathered each trial data into vectors
  profit_trials <- c() #this vector saves profit values
  customers_trials <- c()
  lost_customers_trials <- c()
  storage_cost_trials <- c()
  overstock_trials <- c()
  scooters_sold_trials <- c()
```

**Starting a loop which corresponds to the trials we gave as an input:**
```{r Explaning calculation function 2,echo=TRUE,eval=FALSE}

for (k in 1:trials) {
    #I used a for loop to run trials
    
    customers <- 0  #number of customers
    scooter <- 1    #number of scooters on stock
    scooter_sold <- 0   #number of scooters sold
    storage_cost <- 0   #summary of storage cost
    lost_customer <- 0   #customers that we lost due to lack of scooters
    
    orders <- 0 #this is a counting number used when we make an order and we want to check when it will be delivered
    order_underway_flag <- 0 #this flag is used in order to know if we have made an order already
    #(so we don't make extra orders daily when we are out of stock)
    profit_loss_function <- 0 #This is our PnL function which calculates our total profit
    delivery_day <- 0 #this is an index used in order to organize deliveries
```
  
**Creating a second loop which immitates the number of weeks that we investigate, running poisson process and storing it into an array:**
```{r Explaning calculation function 3,echo=TRUE,eval=FALSE}


  for (j in 1:number_of_weekends) {
    poisson_data <- rpois(7, lambda)
    poisson_array <- array(poisson_data)
    #print(poisson_array)
    
```

***Creating a third loop which immitates the number of days and calculate results using if statements:***
```{r Explaning calculation function 4,echo=TRUE,eval=FALSE}
    

      for (i in 1:7) {
        delivery_day <- delivery_day + 1
        
        #the following if statement is used to add scooters to the inventory when our orders reach maturity(date of delivery)
        if (orders == delivery_day - 5 &
            order_when_out == 1 & order_underway_flag == 1) {
          scooter <- scooter + 1
          orders = 0
          order_underway_flag = 0
        }
        
        #This if statement is used to add scooters on inventory on a Fixed Delivery base
        if (fixed_delivery == 0) {
          #I used this if statement because in case fixed_delivery is disabled
          #and we get an exemption on the next statement *division with 0*
        } else if (delivery_day %% fixed_delivery == 0) {
          #finding mod of delivery day and fixed delivery number
          scooter = scooter + 1
        }
        
        #The following if statement is used to calculate sales and lost customers
        if (poisson_data[i] > 0) {
          customers <- customers + poisson_data[i]
          if (scooter >= poisson_data[i]) {
            scooter <- scooter - poisson_data[i]
            scooter_sold <- scooter_sold + poisson_data[i]
          } else if (scooter < poisson_data[i]) {
            dif <- abs(scooter - poisson_data[i])
            scooter_sold <- scooter_sold + scooter
            scooter <- 0
            lost_customer <- lost_customer + dif
          }
        }
        
        #following if statement is used to create orders when we are out of stock
        if (scooter == 0 &
            order_when_out == 1 & order_underway_flag == 0) {
          orders <- delivery_day
          order_underway_flag = 1
        }
        
        if (scooter > 0) {
          storage_cost <- storage_cost + scooter * 5
        }
      }
    }
```

**Calculating results for each corresponding value and store them in a dataframe, which is then returned as the output of the function.**
Note: While calculating profit I made an assumption that overstock scooters are returned to supplier so the cost of goods sold is not deducted from our profits.

```{r Explaning calculation function 7,echo=TRUE,eval=FALSE}

    profit_loss_function <- scooter_sold * 200 - lost_customer * 100 - storage_cost
    #print(profit_loss_function)
    
    #gathering trial data
    profit_trials[k] <- profit_loss_function
    customers_trials[k] <- customers
    lost_customers_trials[k] <- lost_customer
    storage_cost_trials[k] <- storage_cost
    overstock_trials[k] <- scooter
    scooters_sold_trials[k] <- scooter_sold
    
  }
  
  df <- data.frame(
    profit_trials = profit_trials,
    customers_trials = customers_trials,
    lost_customers_trials = lost_customers_trials,
    storage_cost_trials = storage_cost_trials,
    overstock_trials = overstock_trials,
    scooters_sold_trials = scooters_sold_trials
  )
  
  return(df)
}
```

```{r Running calculation function,echo=FALSE,eval=TRUE}

my_function <- function(number_of_weekends,
                        order_when_out,
                        fixed_delivery) {
  trials <- 10000 #here we set number of trials
  #number_of_weekends <- This is the timeframe we want to check
  
  #order_when_out   : Flag used for orders in case we are out of stock
  #(If value is 1 then whenever our stock is 0 we make an order)
  
  #fixed_delivery: This parameter is used for deliveries that are on a fixed schedule.
  #If it's 0 the parameter is disabled
  
  lambda <- 1 / 7   #Lambda parameter for Poisson
  
  #To get our final result I implemented a for loop of 10.000 trials and gathered each trial data into vectors
  profit_trials <- c() #this vector saves profit values
  customers_trials <- c()
  lost_customers_trials <- c()
  storage_cost_trials <- c()
  overstock_trials <- c()
  scooters_sold_trials <- c()
  
  
  
  for (k in 1:trials) {
    #I used a for loop to run trials
    
    customers <- 0  #number of customers
    scooter <- 1    #number of scooters on stock
    scooter_sold <- 0   #number of scooters sold
    storage_cost <- 0   #summary of storage cost
    lost_customer <- 0   #customers that we lost due to lack of scooters
    
    orders <- 0 #this is a counting number used when we make an order and we want to check when it will be delivered
    order_underway_flag <- 0 #this flag is used in order to know if we have made an order already
    #(so we don't make extra orders daily when we are out of stock)
    profit_loss_function <- 0 #This is our PnL function which calculates our total profit
    delivery_day <- 0 #this is an index used in order to organize deliveries
    
    for (j in 1:number_of_weekends) {
      poisson_data <- rpois(7, lambda)
      poisson_array <- array(poisson_data)
      #print(poisson_array)
      
      for (i in 1:7) {
        delivery_day <- delivery_day + 1
        
        #the following if statement is used to add scooters to the inventory when our orders reach maturity(date of delivery)
        if (orders == delivery_day - 5 &
            order_when_out == 1 & order_underway_flag == 1) {
          scooter <- scooter + 1
          orders = 0
          order_underway_flag = 0
        }
        
        #This if statement is used to add scooters on inventory on a Fixed Delivery base
        if (fixed_delivery == 0) {
          #I used this if statement because in case fixed_delivery is disabled
          #and we get an exemption on the next statement *division with 0*
        } else if (delivery_day %% fixed_delivery == 0) {
          #finding mod of delivery day and fixed delivery number
          scooter = scooter + 1
        }
        
        #The following if statement is used to calculate sales and lost customers
        if (poisson_data[i] > 0) {
          customers <- customers + poisson_data[i]
          if (scooter >= poisson_data[i]) {
            scooter <- scooter - poisson_data[i]
            scooter_sold <- scooter_sold + poisson_data[i]
          } else if (scooter < poisson_data[i]) {
            dif <- abs(scooter - poisson_data[i])
            scooter_sold <- scooter_sold + scooter
            scooter <- 0
            lost_customer <- lost_customer + dif
          }
        }
        
        #following if statement is used to create orders when we are out of stock
        if (scooter == 0 &
            order_when_out == 1 & order_underway_flag == 0) {
          orders <- delivery_day
          order_underway_flag = 1
        }
        
        if (scooter > 0) {
          storage_cost <- storage_cost + scooter * 5
        }
      }
    }
    
    profit_loss_function <- scooter_sold * 200 - lost_customer * 100 - storage_cost
    #print(profit_loss_function)
    
    #gathering trial data
    profit_trials[k] <- profit_loss_function
    customers_trials[k] <- customers
    lost_customers_trials[k] <- lost_customer
    storage_cost_trials[k] <- storage_cost
    overstock_trials[k] <- scooter
    scooters_sold_trials[k] <- scooter_sold
    
  }
  
  df <- data.frame(
    profit_trials = profit_trials,
    customers_trials = customers_trials,
    lost_customers_trials = lost_customers_trials,
    storage_cost_trials = storage_cost_trials,
    overstock_trials = overstock_trials,
    scooters_sold_trials = scooters_sold_trials
  )
  
  return(df)
}

```

***The following is an example of using the function. We input different values for the weeks we want to investigate and save the input based on our delivery terms to simulate Strategy 1: ***
```{r Function usage and storing to vectors example,echo=TRUE,eval=FALSE}
#Strategy 1 metrics
strategy1_monthly<-my_function(4,1,0) #4 weeks
strategy1_querterly<-my_function(13,1,0) #12 weeks
strategy1_yearly<-my_function(52,1,0) #52 weeks
```

```{r Function usage and storing to vectors,echo=FALSE}

#Strategy 1 metrics
strategy1_monthly <- my_function(4, 1, 0) #4 weeks
strategy1_querterly <- my_function(13, 1, 0) #12 weeks
strategy1_yearly <- my_function(52, 1, 0) #52 weeks

#Strategy 2 metrics

strategy2_monthly <- my_function(4, 0, 7) #4 weeks
strategy2_querterly <- my_function(13, 0, 7) #12 weeks
strategy2_yearly <- my_function(52, 0, 7)  #52 weeks

#Strategy 3 metrics

strategy3_monthly_10 <- my_function(4, 1, 10) #4 weeks , fixed delivery every 10 days
strategy3_3month_10 <- my_function(13, 1, 10) #12 weeks, fixed delivery every 10 days
strategy3_yearly_10 <- my_function(52, 1, 10)  #52 weeks, fixed delivery every 10 days

#Testing other delivery schedules
#strategy3_Week1_12<-my_function(1,1,12) #1 week , fixed delivery every 12 days
#strategy3_monthly_12<-my_function(4,1,12) #4 weeks , fixed delivery every 12 days
#strategy3_3month_12<-my_function(13,1,12) #12 weeks, fixed delivery every 12 days
#strategy3_yearly_12<-my_function(52,1,12)  #52 weeks, fixed delivery every 12 days

#strategy3_Week1_15<-my_function(1,1,15) #1 week , fixed delivery every 15 days
#strategy3_monthly_15<-my_function(4,1,15) #4 weeks , fixed delivery every 15 days
#strategy3_3month_15<-my_function(13,1,15) #12 weeks, fixed delivery every 15 days
#strategy3_yearly_15<-my_function(52,1,15)  #52 weeks, fixed delivery every 15 days
```

## Visualization

To visualize results I used the following functions:

**1.get_vectors:** Which gathers trial data for each respective timeline and strategy. It uses as an input the number of weeks and the metric we want to view.For example if we input 4(weeks) and 1(metric) we will get monthly profits for each strategy.
```{r Visualization function 1,echo=TRUE,eval=TRUE}
# This function to select vectors based on weeks and metric
get_vectors <- function(weeks, metric) {
  if (metric == 1) {
    if (weeks == 4) {
      return(list(strategy1_monthly[1], strategy2_monthly[1], strategy3_monthly_10[1]))
    } else if (weeks == 13) {
      return(list(strategy1_querterly[1], strategy2_querterly[1], strategy3_3month_10[1]))
    } else if (weeks == 52) {
      return(list(strategy1_yearly[1], strategy2_yearly[1], strategy3_yearly_10[1]))
    }
  } else if (metric == 2) {
    if (weeks == 4) {
      return(list(strategy1_monthly[3]/strategy1_monthly[2]*100, strategy2_monthly[3]/strategy2_monthly[2]*100, strategy3_monthly_10[3]/strategy3_monthly_10[2]*100))
    } else if (weeks == 13) {
      
      return(list(strategy1_querterly[3]/strategy1_querterly[2]*100, strategy2_querterly[3]/strategy2_querterly[2]*100, strategy3_3month_10[3]/strategy3_3month_10[2]*100))
    } else if (weeks == 52) {
      return(list(strategy1_yearly[3]/strategy1_yearly[2]*100, strategy2_yearly[3]/strategy2_yearly[2]*100, strategy3_yearly_10[3]/strategy3_yearly_10[2]*100))
    }
  } else if (metric == 3) {
    if (weeks == 4) {
      return(list(strategy1_monthly[6]/strategy1_monthly[2]*100, strategy2_monthly[6]/strategy2_monthly[2]*100, strategy3_monthly_10[6]/strategy3_monthly_10[2]*100))
    } else if (weeks == 13) {
      return(list(strategy1_querterly[6]/strategy1_querterly[2]*100, strategy2_querterly[6]/strategy2_querterly[2]*100, strategy3_3month_10[6]/strategy3_3month_10[2]*100))
    } else if (weeks == 52) {
      return(list(strategy1_yearly[6]/strategy1_yearly[2]*100, strategy2_yearly[6]/strategy2_yearly[2]*100, strategy3_yearly_10[6]/strategy3_yearly_10[2]*100))
    }
  } else if (metric == 4) {
    if (weeks == 4) {
      return(list(strategy1_monthly[5], strategy2_monthly[5], strategy3_monthly_10[5]))
    } else if (weeks == 13) {
      return(list(strategy1_querterly[5], strategy2_querterly[5], strategy3_3month_10[5]))
    } else if (weeks == 52) {
      return(list(strategy1_yearly[5], strategy2_yearly[5], strategy3_yearly_10[5]))
    }
  }
}

```

**2. Plot_histogram:** This function is used to return histogram plots with specific parameters.
```{r Visualization function 2,echo=TRUE,eval=TRUE}
# Helper function to plot histograms
# Load the required library
library(ggplot2)

# Function to plot histograms using ggplot with vectors
plot_histograms_ggplot <- function(s1_vector, s2_vector, s3_vector, weeks, metric_name) {
  
  s1_vector <- as.numeric(s1_vector[!is.na(s1_vector)])
  s2_vector <- as.numeric(s2_vector[!is.na(s2_vector)])
  s3_vector <- as.numeric(s3_vector[!is.na(s3_vector)])
  
  p <- ggplot() +
    
    geom_histogram(aes(x = s1_vector, fill = "Strategy 1"), 
                   alpha = 0.8, bins = 30, color = "white", 
                   position = "identity") +
    
    geom_histogram(aes(x = s2_vector, fill = "Strategy 2"), 
                   alpha = 0.9, bins = 30, color = "white", 
                   position = "identity") +
    
    geom_histogram(aes(x = s3_vector, fill = "Strategy 3"), 
                   alpha = 0.9, bins = 30, color = "white", 
                   position = "identity") +
    
    labs(title = paste(weeks, "week", metric_name, "per Strategy"),
         x = metric_name, y = "Frequency",fill = "Strategies") +
    
    scale_fill_manual(values = c("Strategy 1" = "#50336b", 
                                 "Strategy 2" = "#f0baa2", 
                                 "Strategy 3" = "#a69ab5")) +
    
    theme_minimal()
  
  # Print the plot
  return(p)
}
```
**3.Main visualize function:** This function is used to plot histograms.

Metric codes used:

  - 1--> Profit in $
  - 2--> Fraction of served customers (%)
  - 3--> Fraction of lost customers (%)
  - 4--> Overstock
```{r Visualization function 3,echo=TRUE,eval=TRUE}

# Main visualize function
visualize <- function(weeks, metric) {
  # Assuming get_vectors() is a function that retrieves the necessary data
  vectors <- get_vectors(weeks, metric)
  
  #Switch gives us a different name for each input
  if(metric==1){
    metric_name="Profits in $"
  }else if(metric==2){
    metric_name="Customers Lost %"
  }else if(metric==3){
    metric_name="Customers Served %"
  }else if(metric==4){
    metric_name="Overstock"
  } 
 
  # Call the plot function with the retrieved vectors
  plot_histograms_ggplot(vectors[[1]], vectors[[2]], vectors[[3]], weeks, metric_name)
}

```

**4.Barplots:** The last function is called for barplots.

```{r Visualization function 4,echo=TRUE,eval=TRUE}


# Create a bar plots for each value
barplots <- function(metric) {
  # Get the vectors and calculate the mean for each strategy for monthly, quarterly, and yearly time periods
  month_baseline <- c(mean(unlist(get_vectors(4, metric)[[1]]), na.rm = TRUE), #null values removed
                      mean(unlist(get_vectors(4, metric)[[2]]), na.rm = TRUE),
                      mean(unlist(get_vectors(4, metric)[[3]]), na.rm = TRUE))
  
  quarter_baseline <- c(mean(unlist(get_vectors(13, metric)[[1]]), na.rm = TRUE),
                        mean(unlist(get_vectors(13, metric)[[2]]), na.rm = TRUE),
                        mean(unlist(get_vectors(13, metric)[[3]]), na.rm = TRUE))
  
  yearly_baseline <- c(mean(unlist(get_vectors(52, metric)[[1]]), na.rm = TRUE),
                       mean(unlist(get_vectors(52, metric)[[2]]), na.rm = TRUE),
                       mean(unlist(get_vectors(52, metric)[[3]]), na.rm = TRUE))
  
  # Set the metric name based on the provided metric
  if (metric == 1) {
    metric_name <- "Profits in $"
  } else if (metric == 2) {
    metric_name <- "Customers Lost %"
  } else if (metric == 3) {
    metric_name <- "Customers Served %"
  } else if (metric == 4) {
    metric_name <- "Overstock"
  }
  
  # Create separate data for each strategy
  strategy1 <- c(month_baseline[1], quarter_baseline[1], yearly_baseline[1])
  strategy2 <- c(month_baseline[2], quarter_baseline[2], yearly_baseline[2])
  strategy3 <- c(month_baseline[3], quarter_baseline[3], yearly_baseline[3])
  
  # Combine the data for strategies into a matrix
  strategies <- rbind(strategy1, strategy2, strategy3)
  
  # Define colors for each strategy
  color1 <- rgb(80, 51, 107, maxColorValue = 255)  # #50336b
  color2 <- rgb(240, 186, 162, maxColorValue = 255)  # #7b6690
  color3 <- rgb(166, 154, 181, maxColorValue = 255)  # #a69ab5
  
  # Define time periods as x-axis labels
  groups <- c("Monthly", "Quarterly", "Yearly")
  
  # Create the grouped bar plot
  barplot(
    strategies,
    beside = TRUE,             # Creates bars side by side
    names.arg = groups,        # Group labels (time periods)
    col = c(color1, color2, color3),
    main = paste(metric_name, "per Strategy"),
    xlab = "Timeline",
    ylab = metric_name,
    ylim = c(0, max(strategies) * 1.2)
  )
  
  # Add a legend to the plot
  legend(
    "topleft",
    legend = c("Strategy 1", "Strategy 2", "Strategy 3"),
    fill = c(color1, color2, color3),
    bty = "n"           # No box around the legend
  )
}
```
## Analysis

### Profits

**Strategy comparison:**

  - Strategy 1 is highly consistent in generating moderate results as the highest frequency of results     is concentrated just above the $0 profit mark. It shows a significant increase on long-term            baseline. However, as noted before it was outperformed by strategy 3.

  - Strategy 2 struggles to turn a profit in a consistent manner. In more detail this strategy shows a     wider spread of outcomes, with many results concentrated in the negative profit range. This suggests     that Strategy 2 consistently underperforms in comparison to other strategies and is not considered     an optimal solution.
  
  - Strategy 3 shows a more evenly spread distribution across a range of profit outcomes.
  The frequency is distributed more broadly across both positive and negative profits,which indicates    that this strategy carries more risk than strategy 1(which is distributed only in positive outcomes).  However, strategy 3 offers a generous risk/reward trade-off as most outcomes offer very high profits.

```{r Profits,echo=TRUE}
barplots(1)
visualize(4, 1)
visualize(13, 1)
visualize(52, 1)
```

### Fraction of Served and Lost Customers

**Strategy comparison:**

 - Strategy 1 consistently underperforms the other strategies, with a median 20-30% less customers         served approximately.
 
 - Strategy 2 has rapid growth in stock as time progress resulting in very high customer satisfaction       rates.     
 
 - Strategy 3 looks very promising considering customers served in the short-term.However this is            reasonable because at the beginning it has more deliveries than other strategies on the short-term.
```{r Fraction of served customers,echo=TRUE}
library(patchwork)

#I use Align plots because fraction served and fraction lost are supplementary data(they do not require distinct analysis because one is inverse of the other.)
layout(matrix(c(1, 2), 1, 2, byrow = TRUE), widths = c(1, 1))
barplots(3)
barplots(2)

aligned_plots <-function(weeks){ 
  combined_plot=visualize(weeks, 3) / visualize(weeks, 2) + 
  plot_layout(heights = c(2, 2)) +  # Equal heights for both plots
  plot_annotation(theme = theme(plot.margin = margin(10, 10, 30, 10)))
  
  return(combined_plot)
  }


aligned_plots(4) #print histogram for 4 weeks
aligned_plots(13)#print histogram for 13 weeks
aligned_plots(52)#print histogram for 52 weeks

```

## Overstock

**Strategy comparison:**

  - Strategy 1 is highly consistent and shows the lowest number of overstocked scooters in the inventory. Given that its median value is the lowest among all strategies, Strategy 1 appears to be the most optimal solution for minimizing inventory overstock.

  - Strategy 2 shows some volatility results across trials, but consistently yields a significantly higher number of scooters in stock. The median value indicates that this strategy is the least efficient when it comes to managing scooter inventory.

  - Strategy 3 represents a balanced, hybrid approach. While it results in a higher number of scooters in stock compared to Strategy 1, this may contribute to better customer satisfaction, making it a potential trade-off between inventory control and customer service.

```{r Overstock scooters,echo=TRUE}

barplots(4)

visualize(4, 4)
visualize(13, 4)
visualize(52, 4)
```

## Summary

- **Strategy 1 faces significant profit losses primarily due to customer service, as it's products are unavailable when needed. In contrast, Strategy 2 struggles with overstocking,leading to substantially higher storage costs.**

- **In conclusion, Strategy 3, a hybrid model combining fixed scheduled orders every N days with ad-hoc orders when inventory is empty, demonstrates clear superiority in the mid-term (quarterly) to long-term (yearly) performance.**

...
